---
title: "R codes master list"
output: html_document
date: "2023-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


[row, column]


```{r}
new_vector <- c(1,2,3,4,5,6,7,8,9)
```


```{r}
sum_object1_object2 <- sum(object1, object2)
```


```{r}
dataframe <- rbind(dataframe, newrow)
```
add new row to dataframe


my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i

```{r}
library("tidyverse")

library("janitor")

library("skimr")
```


```{r}
read_csv("data/dataset.csv")
```
loads data


```{r}
dataset <- readr::read_csv("data/dataset.csv", na = c("", "-99", "-"))
```


```{r}
dataset <- read_csv(here("lab#", "data#", "dataset.csv"))
```
points to specific location - exactly where to find dataset 

```{r}
janitor::clean_names(dataset)
```
cleans up column names


```{r}
dataset_new <- rename(dataset, newname="oldname", newname2="oldname2")
```


```{r}
dataset <- select_all(dataset, tolower)
```
changes all column names to lower case


```{r}
select_all(dataset, ~str_replace(., " ", "_"))
```
get rid of spaces


```{r}
names(dataset)
```
gives names of columns


```{r}
glimpse(dataset)
```
insight into data


```{r}
str(dataset)
```
structure of data set


```{r}
dim(dataset)
```
dimensions of data set


```{r}
head(dataset)
```
just columns and first few rows


```{r}
tail(dataset, n=)
```
columns and last few rows


```{r}
levels(dataset$column)
```
levels of the column


```{r}
summary(dataset)
```


```{r}
nrow(dataset)
```


```{r}
ncol(dataset)
```


how to find mean: 

```{r}
mean(dataset$nameofcolumnyouwantmeanfrom)
```

OR

```{r}
dataset %>%
  summarise(mean_column=mean(column))
```


```{r}
median(dataset$nameofcolumnyouwantmedianfrom)
```
how to find median


```{r}
sd()
```


```{r}
min()
```


```{r}
max()
```


```{r}
n()
```
returns the length of a column


```{r}
first()
```
returns first value in a column


```{r}
last()
```
returns last value in a column


```{r}
n_distinct()
```
number of distinct values in a column


```{r}
select(column1, column2, column3)
```


```{r}
select(dataset, starts_with(""))
```


```{r}
select(dataset, -column1, -column2)
```
excludes column1 and column2

da


```{r}
ends_with()
```
Select columns that end with a character string  


```{r}
contains()
```
Select columns that contain a character string  


```{r}
matches()
```
Select columns that match a regular expression  


```{r}
one_of()
```
Select columns names that are from a group of names  


```{r}
filter(dataset, col1=="chr", col2==num, col3>=num)
```


```{r}
filter(dataset, column != "")
```
filter rows that do NOT contain " "


```{r}
filter(dataset, column %in% c(1, 2))
```
Filtering multiple values within the same variable requires the `%in%`


```{r}
filter(dataset, between(column, 1, 2))
```
use `between` if you are looking for a range of specific values


```{r}
filter(dataset, near(column, 1, tol = .5))
```
extract observations "near" a certain value but you need to specify a tolerance


& - and


```{r}
filter(condition1 | condition2)
```
will return rows where condition 1 or condition 2 is met


```{r}
filter(xor(condition1, condition2))
```
will return all rows where only one of the conditions is met, and not when both conditions are met


```{r}
arrange()
```
command is a bit like a sort command in excel. Note that the default is ascending order


```{r}
desc()
```
sort in decreasing order


```{r}
mutate()
```
calculate a new column from existing columns


```{r}
select_if(dataset, ~!is.numeric(.))
```
selects columns that are NOT numeric (~!)


```{r}
select_if(dataset, is.numeric)
```
selects columns that are numeric


```{r}
dataset$column <- as.factor(dataset$column)
class(dataset$column)
```
change class of a column to a factor


```{r}
dataset <- 
dataset %>% 
  mutate(across(c(column1, column2, column3, column4), as_factor))
```
change class of multiple columns to factors


```{r}
dataset %>%
 mutate(column_new = ifelse(column == -999.00, NA, column))
```
change -999 to NA


```{r}
summarise(across(contains(" "), mean))
```
several columns contain " " so this creates a table with all the means of colums that contain " "


```{r}
dataset %>%
  group_by() %>%
  summarize(mean_sleep = mean(sleep_total), 
              min_sleep = min(sleep_total),
              max_sleep = max(sleep_total),
              total = n())
```
**can add filter, select, group_by to the summarize/summarise functions to hone in 
use group by with ategorical values
 
              
```{r}
dataset %>%
count(column)
```
how many types (of that column) in dataset


```{r}
count(,sort=T) 
```
sort=T will put in descending order


```{r}
hist(dataset$column)
```
histogram of column data


```{r}
na.rm=TRUE    
is.na()
anyNA
```


```{r}
dataset %>% 
  summarize(number_nas = sum(is.na(dataset)))
```
how many NAs in the dataset


```{r}
dataset %>% 
  summarise_all(~(sum(is.na(.))))
```
where all the NAs are across the columns 


```{r}
dataset %>%
  summarize(across(c(column1, column2, column3), n_distinct))
```
distinct counts over multiple variables


```{r}
dataset %>%
  summarize(across(contains(" "), mean, na.rm=T))
```
get the mean for all variables in dataset that contain " ". 


```{r}
dataset %>%
  summarise_all(n_distinct)
```


```{r}
naniar::miss_var_summary(dataset)
```
nice, clean summary of NA's across the data frame

```{r}
dataset %>%  
  mutate(column_modified = na_if(column, ""))%>% 
  count(column_modified, sort = T)
```

```{r}
getwd()
```
where am I?


```{r}
dataset %>%
  pivot_longer(-column, #will not pivot
               names_to = "new column", #make a new column 
               values_to = "new column", #values moved to
               ) %>%
  filter()%>%
  arrange()
```


```{r}
dataset2 <- 
  dataset %>% 
  pivot_longer(col1:col76, # a range of columns
               names_to = "new col",
               values_to = "new col2",
               values_drop_na = TRUE #this will drop the NA's
               )
```


```{r}
dataset %>% 
  separate(column, into= c("column1", "column2"), sep = "_")
```
 separate columns whose variables look like variable1_variable2 in 1 column
 
 
```{r}
dataset %>%
  unite(newcolumn, col1, col2)
```
 combines col1 and col2


```{r}
dataset %>% 
  transform(column = str_split(column, ";")) %>%
  unnest(column)
```
more than one value or observation in a row separated by ;


```{r}
dataset %>% 
  pivot_wider(names_from = "column", #the observations under key will become new columns
              values_from = "coulmn2")
```


```{r}
dataset %>% 
  filter(column1=="" | column1=="") %>% 
  group_by(column2, column1) %>% 
  summarise(n=n(), .groups='keep')
```



GGPLOT: 

```{r}
labs(title = "",
       x = "",
       y = "")
```
label plot!


```{r}
dataset %>%
  ggplot(aes(x=, y=, fill=)) + 
  geom_boxplot(alpha=.7) 
```
box plot: define x and y. fill in the boxplots by some variable, can be x! alpha changes color saturation


```{r}
dataset%>%
  ggplot(aes(x=, y=, fill=)) + geom_col() 
```
bar plot with y defined 


```{r}
dataset%>%
  ggplot(aes(x=, fill=)) + geom_bar() 
```
bar plot with no defined y

```{r}
dataset %>% 
  ggplot(aes(x=, y=, group=, color=))+
  geom_line()+
  geom_point(shape=)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  labs(title = "",
       x = "",
       fill = "")
```
line graph: shape - canges shape of points on line. theme changes orientation of values on x axis




```{r}
homerange %>% 
  ggplot(aes(x = log10.mass)) +
  geom_histogram(alpha = 0.9, color = "cornsilk1", fill = "plum", bins=25)+
  labs(title = "")
```
histogram
color changes outline, fill changes color inside bins chane how many histogram bars

```{r}
grDevices::colors()
```


```{r}
ls("package:ggthemes")[grepl("theme_", ls("package:ggthemes"))]
```

```{r}
p+theme_classic()+
  theme(axis.text.x = element_text(angle = 60, hjust=1))+
  labs(title = "Observations by Taxon in Homerange Data",
       x = NULL,
       y= "n",
       fill= "Trophic Guild")
```


```{r}
?RColorBrewer
```

```{r}
display.brewer.pal(6,"GnBu")
```

